---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r echo=FALSE}
setwd("C:\\All Work\\Box Sync\\2015\\Analytics Engineer Course\\Reproducible Research\\Course Project")

```
## Loading and preprocessing the data
Data is loaded using `r "read.csv"` and stored in the dataframe `r "mydata"`.
```{r}
library(ggplot2)
mydata = read.csv("activity/activity.csv")
head(mydata)
```

## What is mean total number of steps taken per day?
First we filter out the rows which don't have missing values in the "steps" column. Then the `r "aggregate"` function is used to calculate the average number of steps per day.
```{r}
mydataf = mydata[!is.na(mydata$steps),]
dailystepscount = aggregate(mydataf$steps,by=list(mydataf$date), FUN=sum, na.rm=TRUE)
names(dailystepscount)=c("Date","Total Steps")
head(dailystepscount)
```

The histogram for total number of steps per day is plotted.
```{r}
hist(dailystepscount$`Total Steps`, 
     breaks=20, 
     col="green",
     xlab="Total Steps",
     main="Histogram of Total Steps")
```

Mean and median of the total steps per day:
```{r}
mean(dailystepscount$`Total Steps`)
median(dailystepscount$`Total Steps`)
```

## What is the average daily activity pattern?

```{r}
avgstepts = aggregate(mydata$steps,by=list(mydata$interval), FUN=mean, na.rm=TRUE)
names(avgstepts) = c("interval","Average Steps")
ggplot(avgstepts, aes(x=interval, y=`Average Steps`))+
  geom_line(col="red")
```

Interval with Maximum number of average steps:
```{r}
avgstepts[which.max(avgstepts$`Average Steps`),]
```

#Inputting Missing Values
Total number of rows with missing values:
```{r}
nrow(mydata[is.na(mydata$steps),])
```
Missing values are filled by using the **Mean for the Interval**
```{r}
missingvals <- mydata[is.na(mydata$steps),]
nomissingvals <- mydata[!is.na(mydata$steps),]
names(avgstepts)=c('interval','steps')
mergevals <- merge(avgstepts, missingvals, by='interval')
filledmissingvals <- mergevals[,c('steps.x','date','interval')]
names(filledmissingvals)[1]='steps'
```
New data set with the filled missing values:
```{r}
mydatafilled <- rbind(nomissingvals,filledmissingvals)
nrow(is.na(mydatafilled$steps))
```

Daily total number of steps for the new data set:
```{r}
dailystepscount = aggregate(mydatafilled$steps,by=list(mydatafilled$date), FUN=sum, na.rm=TRUE)
names(dailystepscount)=c("Date","Total Steps")
head(dailystepscount)
```

Histogram of the new data set:
```{r}
hist(dailystepscount$`Total Steps`, 
     breaks=20, 
     col="blue",
     xlab="Total Steps",
     main="Histogram of Total Steps")

mean(dailystepscount$`Total Steps`)
median(dailystepscount$`Total Steps`)
```

## Are there differences in activity patterns between weekdays and weekends?
```{r}
mydatafilled$date <- as.Date(mydatafilled$date)
mydatafilled$day = weekdays(mydatafilled$date)
mydatafilled$daytype = ''
mydatafilled[mydatafilled$day %in% c("Saturday","Sunday"),'daytype']="Weekend"
mydatafilled[!mydatafilled$day %in% c("Saturday","Sunday"),'daytype']="Weekday"
mydatafilled$daytype <- factor(mydatafilled$daytype)
avgstepsdaytype = aggregate(mydatafilled$steps, by=list( mydatafilled$interval, mydatafilled$daytype), FUN=mean)
names(avgstepsdaytype)=c("Interval","DayType","Average Steps")
ggplot(data=avgstepsdaytype, aes(x=Interval, y=`Average Steps`))+
  geom_line(color="blue")+
  facet_grid(.~ DayType)
```

It can be observed that the movemnt i.e. the number of steps are more uniform across the day during the weekends. 
